---
title: Plot Kozak strength of initiation mutations based on Dvir et al. PNAS 2013
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.time(), '%d %B, %Y')`"

output: 
  github_document:
    toc: 2
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```

# Load libraries that we use repeatedly

```{r}
library(tidyverse)
library(plyranges)
library(Biostrings)
library(rasilabRtemplates)
```

# Download the Table S1 from Dvir et al. that has the protein abundance for 2041 sequence variants.

```{bash}
cd ../tables/
wget --no-clobber http://www.pnas.org/lookup/suppl/doi:10.1073/pnas.1222534110/-/DCSupplemental/sd01.xlsx
```

Read the data from Dvir et al.

```{r}
dvir_data <- readxl::read_excel("../tables/sd01.xlsx") %>% 
  print()
```

The `sequence_variant` column has the 10 nt preceding the start codon.
Average protein abundance over all positions except the last 4 nt.
Then arrange from highest to lowest expression.

```{r}
dvir_data <- dvir_data %>% 
  mutate(kozak = str_extract(sequence_variant, "\\w{4}$")) %>% 
  group_by(kozak) %>% 
  summarize(exp = mean(mean_protein_abundance), 
            se_exp = sd(mean_protein_abundance)/ sqrt(dplyr::n() - 1)) %>% 
  arrange(desc(exp)) %>% 
  print()
```

# Plot Kozak strength vs Kozak sequence for paper

```{r, fig.width=1.5, fig.height=1.5}
# intiiation sites are arranged in this order
initiationmutation_order <- seq(1,8)
names(initiationmutation_order) <- toupper(c( 'ctgc', 'ccgc', 
                              'acgc', 'ccga', 'ccac', 'ccaa', 'caaa', 'aaaa'))
plot_data <- dvir_data %>% 
  filter(kozak %in% names(initiationmutation_order)) %>% 
  mutate(kozak = fct_relevel(kozak, names(initiationmutation_order)))
  
plot_data %>% 
  ggplot(aes(x = kozak, y = exp, 
             ymin = exp - se_exp, ymax = exp + se_exp, group = NA)) +
  geom_point(size = 1, height = 0, width = 0.1, alpha = 0.5) +
  geom_line() +
  geom_errorbar(width = 0.5) +
  labs(y = 'fluorescence (a.u.)',
       x = '-4 to -1 nt from ATG') +
  theme(legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  scale_y_continuous(breaks = seq(3,15, 3))

ggsave("../figures/kozak_mutations_compare_dvir.pdf")
```

# Source data for S1 Fig panel C

```{r}
plot_data %>% 
  arrange(kozak) %>% 
  select(kozak, exp, se_exp) %>% 
  mutate_if(is.numeric, funs(round(., 3))) %>% 
  knitr::kable()
```


# Session Info

```{r}
sessionInfo()
```

