---
title: "Fluorescence of PGK1 WT reporters with endogenous stall inserts and in WT and ASC1Δ strains"
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Import libraries and analysis-specific parameters

```{r}
# standard analysis and plotting functions, includes dplyr, ggplot2 
library(tidyverse)
# loads lab default ggplot2 theme and provides color-blind friendly palette
library(rasilabRtemplates)
# standard error
library(plotrix)
# to tidy up statistical test results
library(broom)

# this folder contains the data and annotations
fcs_file_folder <- "../../data/flow/endogenous_gene_stall/"

# These are the PAR32 plasmid 1 that had a deletion of an AAG
bad_wells <- tibble(strain = c("schp1499", "schp1491"))
```

```{r, eval = FALSE, include = FALSE}
# Combine raw .fcs files into a single dataframe for manuscript
# This chunk is for preprocessing and is not evaluated (`eval = FALSE`).

raw_data <- fcs_file_folder %>% 
  # get all fcs files
  list.files(full.names = T, pattern = '.fcs$') %>% 
  enframe("sno", "filename") %>% 
  # get data from each file
  mutate(data = map(filename, . %>% 
                      flowCore::read.FCS(transformation = F, alter.names = T) %>% 
                      flowCore::exprs() %>% 
                      as_tibble())) %>% 
  # extract plate and well
  mutate(plate = as.integer(str_extract(filename, "(?<=Specimen_)[:digit:]{3}"))) %>%
  mutate(well = str_extract(filename, "(?<=Specimen_[:digit:]{3}_)[:alnum:]+")) %>%
  # get rid of unwanted columns
  select(-sno, -filename) %>% 
  unnest() %>% 
  # round nicely for storing
  mutate_at(vars(matches("Time")), funs(round(., 3))) %>% 
  mutate_at(vars(-matches("Time|well")), funs(as.integer)) %>% 
  print()
```

```{r, eval = FALSE, include = FALSE}
# Clean annotations file for manuscript
# This chunk is for preprocessing and is not evaluated (`eval = FALSE`).

annotations <- paste0(fcs_file_folder, '/sampleannotations.tsv') %>% 
                        read_tsv(col_types = list('plate' = col_integer())) %>% 
  # replace all 'na' ny NA
  mutate_all(funs(ifelse(. == 'na', NA, .))) %>%
  # convert to lower case
  mutate(strain = tolower(strain)) %>% 
  # these are genes of interest or the standard control strains
  select(plate, well, strain, gene, stall, kogene, replicate) %>%
  write_tsv(paste0(fcs_file_folder, '/annotations.tsv')) %>%
  print()
```

```{r, eval = FALSE, include = FALSE}
# Keep only the data that we need for manuscript
# This chunk is for preprocessing and is not evaluated (`eval = FALSE`).
raw_data %>% 
  right_join(select(annotations, plate, well)) %>% 
  write_tsv(paste0(fcs_file_folder, '/data.tsv.xz'))
```

# Read data

```{r}
flowdata  <- read_tsv(paste0(fcs_file_folder, '/data.tsv.xz')) %>% 
  print()
```

# Read annotations

```{r}
annotations  <- read_tsv(paste0(fcs_file_folder, '/annotations.tsv')) %>% 
  print()
```

# Rename and calculate average values of fluorescence channels in each well

```{r}
by_file <- flowdata  %>% 
  # group by each plate and well
  group_by(plate, well) %>% 
  select(FITC.A, PE.Texas.Red.A) %>% 
  # calculate mean
  summarise_all(mean) %>% 
  # rename
  rename('yfp' = FITC.A, 'rfp' = PE.Texas.Red.A) %>% 
  # join annotations
  left_join(annotations, by = c('plate', 'well')) %>% 
  print()
```

# Calculate mean and standard error over replicates

```{r}
avg_data  <- by_file %>% 
  anti_join(bad_wells) %>%
  # strain is used to get replicates
  group_by(strain) %>% 
  # calculate mean and std.err
  mutate(mean_yfp = mean(yfp), 
         mean_rfp = mean(rfp)) %>% 
  ungroup() %>% 
  mutate(yfp = yfp - mean_yfp[strain == "schp15" & replicate == 1], 
         rfp = rfp - mean_rfp[strain == "by4741" & replicate == 1]) %>% 
  mutate(yfp_rfp_ratio = yfp / rfp) %>% 
  # calculate mean and standard error
  group_by(strain) %>% 
  mutate(mean_yfp = mean(yfp), 
         mean_rfp = mean(rfp), 
         mean_ratio = mean(yfp_rfp_ratio), 
         se_yfp = std.error(yfp), 
         se_rfp = std.error(rfp),
         se_ratio = std.error(yfp_rfp_ratio),
         n = n()) %>% 
  ungroup()

normalization <- avg_data %>% 
  filter(strain == "schp19")
```

# Plot and tabulate background subtracted ratio of YFP/RFP between ASC1KO and WT strains (used for calculating P-value)

```{r, fig.width=2, fig.height=1.6}
plot_data <- avg_data %>% 
  # mutate(plasmid_replicate = as.factor(plasmid_replicate)) %>% 
  filter(gene == "PGK1") %>% 
  group_by(strain) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(stall) %>% 
  mutate(mean_asc1_wt_ratio = mean_ratio / mean_ratio[is.na(kogene)] * 100) %>% 
  mutate(se_asc1_wt_ratio = (se_ratio / mean_ratio + 
                               se_ratio[is.na(kogene)] / mean_ratio[is.na(kogene)]) * 
           mean_asc1_wt_ratio) %>% 
  ungroup() %>% 
  mutate(stall = if_else(is.na(stall), "No stall", as.character(stall))) %>% 
  mutate(stall = fct_reorder(stall, -mean_asc1_wt_ratio)) %>% 
  filter(!is.na(kogene)) %>% 
  print()

plot_data %>%
  ggplot(aes(x = stall, y = mean_asc1_wt_ratio,
             ymin = mean_asc1_wt_ratio - se_asc1_wt_ratio,
             ymax = mean_asc1_wt_ratio + se_asc1_wt_ratio,
             group = NA)) +
  # geom_point(size = 1.5) +
  geom_col() +
  geom_errorbar(width = 0.25) +
  labs(y = 'ASC1 / WT \n protein level ratio (%)',
       x = 'Endogenous stall') +
  scale_color_manual(values = cbPalette) +
  scale_y_continuous(limits = c(0, NA)) +
  theme(legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6)
        )
```

# Plot and tabulate background subtracted YFP/RFP ratio for endogenous stalls

```{r, fig.width=1.8, fig.height=1.6}
plot_data <- avg_data %>% 
  mutate(mean_ratio = mean_ratio / normalization[[1, "mean_ratio"]]) %>% 
  mutate(se_ratio = se_ratio / normalization[[1, "mean_ratio"]]) %>% 
  filter(gene == "PGK1") %>% 
  group_by(strain) %>% 
  slice(1) %>% 
  ungroup() %>% 
  group_by(stall) %>% 
  mutate(mean_asc1_wt_ratio = mean_ratio / mean_ratio[is.na(kogene)] * 100) %>% 
  ungroup() %>% 
  mutate(kogene = fct_rev(if_else(is.na(kogene), "WT", as.character(kogene)))) %>% 
  mutate(stall = if_else(is.na(stall), "No stall", as.character(stall))) %>% 
  mutate(stall = fct_reorder(stall, - mean_asc1_wt_ratio)) %>% 
  print()

plot_data %>%
  ggplot(aes(x = stall, y = mean_ratio,
             ymin = mean_ratio - se_ratio,
             ymax = mean_ratio + se_ratio,
             color = kogene,
             group = kogene)) +
  geom_errorbar(width = 0.5, color = "black") +
  geom_point(size = 2) +
  geom_line() +
  labs(y = 'Protein level (a.u.)',
       x = 'Endogenous gene stall') +
  scale_color_manual(values = cbPalette[2:3]) +
  theme(legend.title = element_text(size = 8),
        axis.text.x = element_text(angle = 45, hjust = 1, size = 6),
        legend.position = "None"
        ) +
  scale_y_continuous(trans = "log2",
                     limits = c(8, 34),
                     breaks = c(8, 16, 32),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)))

ggsave("figures/endogenous_stalls.pdf")
```

# Source data for Fig. 6B, right panel

```{r}
avg_data %>%
  mutate(mean_ratio = mean_ratio / normalization[[1, "mean_ratio"]]) %>%
  mutate(se_ratio = se_ratio / normalization[[1, "mean_ratio"]]) %>%
  filter(gene == "PGK1" & replicate == 1) %>%
  select(kogene, stall, mean_ratio, se_ratio) %>%
  arrange(stall, kogene) %>% 
  knitr::kable()
```

We will normalize the YFP RFP ratio for each strain by the mean no stall value for the WT or ASC1 strain. Then we can compare the ASC1 and WT strain for significance for each stall.

```{r}
plot_data <- avg_data %>% 
  # mutate(plasmid_replicate = as.factor(plasmid_replicate)) %>% 
  filter(gene == "PGK1") %>% 
  group_by(kogene, replicate) %>% 
  mutate(norm_yfp_rfp_ratio = yfp_rfp_ratio / yfp_rfp_ratio[is.na(stall)]) %>% 
  ungroup() %>% 
  select(kogene, stall, replicate, norm_yfp_rfp_ratio) %>% 
  print()
```

# Display t.test P-value for difference in YFP/RFP ratio between WT and ASC1 strains for each stall

```{r}
plot_data %>% 
  mutate(kogene = if_else(is.na(kogene), "WT", as.character(kogene))) %>%
  # get rid of the no stall control since this has P-value of 1 by definition
  filter(!is.na(stall)) %>%
  group_by(stall) %>%
  nest() %>%
  mutate(t.test = map(data, function(df) tidy(t.test(norm_yfp_rfp_ratio ~ kogene, df, alternative = "greater")))) %>%
  unnest(t.test) %>%
  select(stall, p.value, statistic) %>%
  arrange(p.value) %>%
  knitr::kable()
```