---
title: "Analyze simulation results"
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

## Load libraries
```{r}
library(tidyverse)
library(rasilabRtemplates)
# disable scientific notation
options(scipen=999)

model_names <- c(
  "hit3" = "CAT",
  "hit5" = "CSAT",
  "simple" = "SAT",
  "trafficjam" = "TJ"
 )
```

# Read protein count data
```{r}
psr_data <- read_tsv("tables/psr_stats.tsv") %>% 
  print()
```

# Read collision data

```{r}
collision_data <- read_tsv("tables/collision_stats.tsv") %>% 
  print()
```

## Read simulation parameters
```{r}
annotations  <- read_tsv('sim.params.tsv', 
                         col_types = cols(x_stall = col_character(),
                                          k_elong_stall = col_character())) %>%
  mutate(preterm_intact_model = if_else(preterm_intact_rate == 0, 
                                        "trafficjam", 
                                        preterm_intact_model)) %>% 
  rename(sim_id = X1) %>%
  mutate(k_elong_stall = str_split(k_elong_stall, ",")) %>%
  mutate(k_elong_stall = map(k_elong_stall, as.numeric)) %>%
  mutate(k_elong_stall = map(k_elong_stall, function(x) unique(x))) %>%
  unnest() %>%
  mutate(x_stall = stringr::str_split(x_stall, ',')) %>%
  mutate(k_stall = k_elong_stall / as.numeric(n_stall)) %>%
  mutate(n_stall = factor(n_stall)) %>%
  select(sim_id, k_init, k_elong_stall, k_stall, x_stall, n_stall,
         preterm_intact_model, preterm_intact_rate) %>%
  print()
```

## Combine all data into a single table
```{r}
data <- annotations %>% 
  left_join(psr_data, by = "sim_id") %>% 
  left_join(collision_data, by = "sim_id") %>% 
  print()
```

## How does PSR vary as a function of initiation rate in all models with low stall elongation rate and medium preterm_intact_rate?


```{r psr_no_quality_control, fig.width=2.4, fig.height=1.5}
plot_data <- data %>% 
  mutate(k_stall = round(k_stall, 1)) %>%
  filter(k_stall == 0.1) %>%
  filter(((preterm_intact_rate == 0 | preterm_intact_rate == 0.02) 
          & preterm_intact_model == "simple") |
        ((preterm_intact_rate == 0 | preterm_intact_rate == 1) 
                & preterm_intact_model != "simple")) %>% 
  mutate(model = forcats::fct_rev(model_names[preterm_intact_model]))

plot_data %>% 
  ggplot(aes(x = k_init, y = psr, shape = model, fill = model, color = model)) +
  geom_point(size = 1.5) + geom_line(size = 0.5) +
  scale_x_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8, 2^0)) +
  scale_y_continuous(breaks = seq(0,0.09, 0.03)) +
  scale_fill_manual(values = cbPalette) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(25, 24, 21, 23)) +                                                  
  labs(x = "initiation rate (s-1)", y = "protein synthesis rate (s-1)",
       fill = "", shape = "", color = "") +
  theme(legend.key.height = unit(0.2, "in"))                     
                         
ggsave('figures/psr_all_models_medium_stall_medium_pretermintact.pdf') 
```

# Change in maximal protein synthesis as a function of stall elongation rate in TJ model
```{r, fig.width=2.25, fig.height=1.45}
plot_data <- data %>% 
  filter(preterm_intact_rate == 0)

plot_data %>% 
  ggplot(aes(x = k_init, y = psr, color = as.factor(k_stall), shape = as.factor(k_stall))) +
  geom_point(size = 1.5) + geom_line(size = 0.5) +
  scale_x_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8.5, 2^0)) +
  scale_y_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8.5, 2^0)) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(19,17,15)) +
  labs(x = "initiation rate (s-1)", y = "protein synthesis rate (s-1)",
       shape = "kstall (s-1)", color = "kstall (s-1)") +
  theme(legend.key.height = unit(0.2, "in")) +
  geom_vline(aes(xintercept = k_stall, color = as.factor(k_stall)), 
             show.legend = F, linetype = "dotted")
                         
ggsave('figures/psr_tj_model_vary_stall_strength.pdf') 

```

# Change in protein synthesis as a function of abort rate in SAT model

```{r, fig.width=2.25, fig.height=1.45}
plot_data <- data %>% 
  mutate(k_stall = round(k_stall, 1)) %>%
  filter(k_stall == 0.1) %>% 
  filter(preterm_intact_rate < 0.1) %>% 
  mutate(preterm_intact_model = if_else(preterm_intact_rate == 0, "simple",
                                        preterm_intact_model)) %>% 
  filter(preterm_intact_model == "simple") %>% 
  mutate(model = forcats::fct_rev(model_names[preterm_intact_model]))

plot_data %>% 
  ggplot(aes(x = k_init, y = psr, color = as.factor(preterm_intact_rate), 
             shape = as.factor(preterm_intact_rate))) +
  geom_point(size = 1.5) + geom_line(size = 0.5) +
  scale_x_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8.5, 2^0)) +
  scale_y_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x))
                     ) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(19,17,15,18,16)) +
  labs(x = "initiation rate (s-1)", y = "protein synthesis rate (s-1)",
       shape = "kabort (s-1)", color = "kabort (s-1)") +
  theme(legend.key.height = unit(0.2, "in")) +
                         
ggsave('figures/psr_sat_model_vary_abort_rate.pdf', width = 2.25, height = 1.45) 
```

# Change in protein synthesis as a function of stall elongation rate in SAT model

```{r, fig.width=2.25, fig.height=1.45}
plot_data <- data %>% 
  filter(preterm_intact_rate == 0.02) %>% 
  filter(preterm_intact_model == "simple") %>% 
  mutate(model = forcats::fct_rev(model_names[preterm_intact_model]))

plot_data %>% 
  ggplot(aes(x = k_init, y = psr, color = as.factor(k_stall), 
             shape = as.factor(k_stall))) +
  geom_point(size = 1.5) + geom_line(size = 0.5) +
  scale_x_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8.5, 2^0)) +
  scale_y_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x))
                     ) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(19,17,15,18,16)) +
  labs(x = "initiation rate (s-1)", y = "protein synthesis rate (s-1)",
       shape = "kstall (s-1)", color = "kstall (s-1)") +
  theme(legend.key.height = unit(0.2, "in")) +
  geom_vline(aes(xintercept = k_stall, color = as.factor(k_stall)), 
             show.legend = F, linetype = "dotted")
                         
ggsave('figures/psr_sat_model_vary_stall_rate.pdf', width = 2.25, height = 1.45) 
```


# Change in protein synthesis as a function of stall elongation rate in CSAT model

```{r, fig.width=2.25, fig.height=1.45}
plot_data <- data %>% 
  filter(preterm_intact_rate == 1) %>% 
  filter(preterm_intact_model == "hit5") %>% 
  mutate(model = forcats::fct_rev(model_names[preterm_intact_model]))

plot_data %>% 
  ggplot(aes(x = k_init, y = psr, color = as.factor(k_stall), 
             shape = as.factor(k_stall))) +
  geom_point(size = 1.5) + geom_line(size = 0.5) +
  scale_x_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x),
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     limits = c(2^-8.5, 2^0)) +
  scale_y_continuous(trans = "log2",
                     breaks = scales::trans_breaks("log2", function(x) 2^x, n = 4),
                     labels = scales::trans_format("log2", scales::math_format(2^.x))
                     ) +
  scale_color_manual(values = cbPalette) +
  scale_shape_manual(values = c(19,17,15,18,16)) +
  labs(x = "initiation rate (s-1)", y = "protein synthesis rate (s-1)",
       shape = "kstall (s-1)", color = "kstall (s-1)") +
  theme(legend.key.height = unit(0.2, "in")) +
  geom_vline(aes(xintercept = k_stall, color = as.factor(k_stall)), 
             show.legend = F, linetype = "dotted")
                         
ggsave('figures/psr_csat_model_vary_stall_rate.pdf', width = 2.25, height = 1.45) 
```

# Collision rate as a function of initiation rate 

```{r, fig.width=1.4, fig.height=2}
plot_data <- data %>% 
  mutate(k_stall = round(k_stall, 1)) %>%
  filter(k_stall == 0.1) %>%
  filter((preterm_intact_rate == 0.05 & preterm_intact_model == "simple") |
        (preterm_intact_rate == 1 & preterm_intact_model == "hit5")) %>% 
  mutate(model = model_names[preterm_intact_model])

plot_data %>%
  ggplot(aes(x = k_init, y = collision_freq, color = model, shape = model)) +
  geom_point(size = 1.5) +
  geom_line() +
  scale_x_continuous(trans = "log2",
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     breaks = 2^(seq(-8,0,4))) +
  # scale_y_continuous(trans = "log2",
  #                    labels = scales::trans_format("log2", scales::math_format(2^.x))) +
  scale_color_manual(values = cbPalette[c(3,2)]) +
  scale_shape_manual(values = c(19, 17,  16)) +
  labs(x = "initiation rate (s-1)", y = "collision frequency (s-1)", color = "", shape = "") +
  theme(legend.position = "top")

ggsave("figures/collision_rate_vs_initiation_rate.pdf")
```

# Same as above but with log Y-scale

```{r, fig.width=0.8, fig.height=0.8}
plot_data <- data %>% 
  mutate(k_stall = round(k_stall, 1)) %>%
  filter(k_stall == 0.1) %>%
  filter((preterm_intact_rate == 0.05 & preterm_intact_model == "simple") |
        (preterm_intact_rate == 1 & preterm_intact_model == "hit5")) %>% 
  mutate(model = model_names[preterm_intact_model])

plot_data %>%
  ggplot(aes(x = k_init, y = collision_freq, color = model, shape = model)) +
  geom_point(size = 1, show.legend = F) +
  geom_line(show.legend = F) +
  theme(axis.text = element_text(size = 6)) +
  scale_x_continuous(trans = "log2",
                     labels = scales::trans_format("log2", scales::math_format(2^.x)),
                     breaks = 2^(seq(-8,0,4))) +
  scale_y_continuous(trans = "log2",
                     labels = scales::trans_format("log2", scales::math_format(2^.x))) +
  scale_color_manual(values = cbPalette[c(3,2)]) +
  scale_shape_manual(values = c(19, 17,  16)) +
  labs(x = "", y = "", color = "", shape = "") +

ggsave("figures/collision_rate_vs_initiation_rate_log_scale.pdf", width=0.8, height=0.8)
```
