---
title: "Analyze initiation codon pair mRNA levels"
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, message=F)
```

# Load libraries and define analysis-specific parameters
```{r}
library(glue)
library(tidyverse)
library(rasilabRtemplates)

codonnames <- c(
  "AAG" = "10xAAG",
  "AGA" = "10xAGA",
  "CCG" = "8xCCG",
  "CCA" = "8xCCA",
  "AGT" = "PGK1_wt",
  "TCC" = "PGK1_5xAGA5")

initiationmutation_order <- seq(1,9)
names(initiationmutation_order) <- c('CTG', 'CTGC', 'CCGC', 
                              'ACGC', 'CCGA', 'CCAC', 'CCAA', 'CAAA', 'AAAA')
```

# Read barcode and strain annotations

```{r}
r2_annotations <- read_tsv("../../data/htseq/r2_barcode_annotations.tsv") %>% 
  print()
strain_annotations <- read_tsv("../../data/htseq/strain_annotations.tsv") %>% 
  print()
barcode_annotations <- read_tsv("../../data/htseq/barcode_annotations.tsv") %>% 
  print()
```

```{r}
strain_barcode_annotations <- read_tsv("../../data/htseq/strain_barcode_annotations.tsv") %>% 
  # get rid of plasmid well and Description, we don't need it
  select(-Description, -well) %>% 
  # combine plate and well location into 1
  mutate(oligo1 = paste0(plate1, "_", well1), 
         oligo2 = paste0(plate2, "_", well2),
         oligo3 = paste0(plate3, "_", well3), 
         oligo4 = paste0(plate4, "_", well4)) %>% 
  # get ride of plate and well columns
  select(-matches("^plate|^well")) %>% 
  # gather all oligos into single well
  gather(oligo, location, -plasmid) %>% 
  # separate again into plate and well
  separate(location, c("plate", "well")) %>% 
  select(-oligo) %>% 
  # clean up A01 to A1, B02 to B2 etc.
  mutate(well = str_replace(well, "([A-Z])0(\\d)", "\\1\\2")) %>% 
  # remove empty wells  
  filter(well != "NA") %>% 
  print()
```

# Read count data and join with barcode and strain annotations

```{r}
count_data <- list.files("tables/", pattern = "barcode_counts.tsv",
                         full.names = T) %>% 
  enframe("sno", "file") %>% 
  # extract R2 barcode
  mutate(barcode = str_extract(file, "(?<=/)[ACTG]{6}")) %>% 
  # read data
  mutate(data = map(file, read_tsv)) %>% 
  # expand to a single dataframe
  unnest() %>% 
  # join by R2 barcode
  left_join(r2_annotations, by = "barcode") %>% 
  # keep only columns of interest
  select(-sno, -file, -index, -barcode, -strain_num, -primer) %>%
  rename(barcode = r1_barcode) %>% 
  left_join(barcode_annotations, by = "barcode") %>% 
  # keep only columns of interest
  select(-primer_name) %>% 
  left_join(strain_barcode_annotations, by = c("plate", "well")) %>% 
  # keep only columns of interest
  select(-plate, -well) %>% 
  left_join(strain_annotations, by = "plasmid") %>% 
  # keep only columns of interest
  select(-row) %>% 
  arrange(count) %>%
  print()
```

# Calculate log2 fold mRNA levels median normalized within each initiation set
```{r}
lfc_data <- count_data %>% 
  # get rid of low counts
  filter(count >= 100) %>% 
  # spread cdna and gdna counts for each barcode to adjacent columns
  spread(type, count) %>%
  # calculate log2 fold change cdna / gdna
  mutate(lfc = log2(cdna)-log2(gdna)) %>%
  # get rid of NA
  filter(!is.na(lfc)) %>%
  # median normalize each group
  group_by(label, strain, init) %>%
  summarize(mean_lfc = mean(lfc), se_lfc = var(lfc) / sqrt(n() - 1), n = n()) %>%
  ungroup() %>%
  group_by(label, strain) %>%
  mutate(mean_lfc = mean_lfc - median(mean_lfc)) %>%
  ungroup() %>%
  # arrange init_mutation in correct_order
  mutate(init = fct_reorder(init, initiationmutation_order[init])) %>%
  print()
```


# Plot mRNA level of PGK1-YFP with different codons, wild-type cells

```{r, fig.width=4, fig.height=1.6}
plot_data <- lfc_data %>% 
  filter(strain == "scHP15" & !str_detect(label, "pgk1")) %>% 
  mutate(codon_group = substr(label, 1, nchar(label)-3)) %>% 
  mutate(stall = if_else(str_detect(label, "aag|ccg"), "yes", "no")) 

plot_data %>% 
  # exclude the mutated start codon since this will be shown in previous panel
  filter(init != "CTG") %>% 
  ggplot(aes(x = init, y = mean_lfc, group = label, shape = stall, color = stall)) +
  facet_wrap(~ codon_group, ncol = 2, scales = "free_y") +
  geom_errorbar(aes(ymin = mean_lfc - se_lfc, ymax = mean_lfc + se_lfc),
                width = 0.5, color = "black") +
  geom_line() +
  geom_point(size = 2) +
  scale_color_manual(values = cbPalette) +
  # scale_shape_manual(values = c(21, 24, 22, 25)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  scale_y_continuous(limits = c(-1.1, +0.5)) +
  labs(x = "-4 to -1 nt from ATG", y = "mRNA level (log2, a.u.)")

ggsave("figures/mrna_level_wt_4_codons.pdf")

```

# Source data for Fig. 2C

```{r}
plot_data %>% 
  # exclude the mutated start codon since this will be shown in previous panel
  filter(init != "CTG") %>% 
  select(label, init, mean_lfc, se_lfc, n) %>% 
  arrange(label, init) %>% 
  mutate_if(is.numeric, funs(round(., 3))) %>% 
  knitr::kable()
```

# Plot mRNA level of PGK1-YFP, no insert

```{r, fig.width=1.6, fig.height=1.6}
plot_data <- lfc_data %>% 
  filter(strain == "scHP15" & str_detect(label, "pgk1, no insert"))

plot_data %>% 
  ggplot(aes(x = init, y = mean_lfc, group = label)) +
  geom_errorbar(aes(ymin = mean_lfc - se_lfc, ymax = mean_lfc + se_lfc),
                width = 0.5, color = "black") +
  geom_line() +
  geom_point(size = 2) +
  scale_color_manual(values = cbPalette) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  # scale_y_continuous(limits = c(-1.1, +0.5)) +
  labs(x = "-4 to -1 nt from ATG", y = "mRNA level (log2, a.u.)")

ggsave("figures/mrna_level_wt_pgk1_no_insert.pdf")
```

# Source data for Fig. 2B

```{r}
plot_data %>% 
  select(init, mean_lfc, se_lfc, n) %>% 
  arrange(init) %>% 
  mutate_if(is.numeric, funs(round(., 3))) %>% 
  knitr::kable()
```

# Plot mRNA levels of  KO codon mutants for paper

```{r, fig.width=5.5, fig.height=1.6}
plot_data <- lfc_data %>% 
  filter(str_detect(strain, "Δ") & !str_detect(label, "pgk1")) %>% 
  mutate(codon_group = substr(label, 1, 1)) %>% 
  filter(codon_group == "8") %>% 
  mutate(stall = if_else(str_detect(label, "aag|ccg"), "yes", "no")) %>% 
  mutate(strain = str_replace(strain, "(.+)(.)", "\\2\\1")) %>% 
  mutate(strain = fct_relevel(strain, "ΔLTN1", "ΔDOM34", "ΔHEL2", "ΔASC1"))

plot_data %>% 
  # exclude the mutated start codon since this will be shown in previous panel
  filter(init != "CTG") %>% 
  ggplot(aes(x = init, y = mean_lfc, group = label, shape = stall, color = stall)) +
  facet_wrap(~ strain, ncol = 4, scales = "fixed") +
  geom_errorbar(aes(ymin = mean_lfc - se_lfc, ymax = mean_lfc + se_lfc),
                width = 0.5, color = "black") +
  geom_line() +
  geom_point(size = 2) +
  scale_color_manual(values = cbPalette) +
  # scale_shape_manual(values = c(21, 24, 22, 25)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) +
  scale_y_continuous(limits = c(-1.1, +0.5)) +
  labs(x = "-4 to -1 nt from ATG", y = "mRNA level (log2, a.u.)")

ggsave("figures/mrna_level_ko_2_codons.pdf")
```

# Source data for Fig. 5C

```{r}
plot_data %>% 
  select(strain, label, init, mean_lfc, se_lfc, n) %>% 
  arrange(strain, label, init) %>% 
  mutate_if(is.numeric, funs(round(., 3))) %>% 
  knitr::kable()
```
